---
title: "Auston Matthews"
format: html
editor: visual
---

# Part 01 - Exploring data

## 1.1 - Loading packages

```{r}
library(tidyverse)
library(reshape2)
library(sqldf)
library(patchwork)
```

## 1.2 - Importing data

```{r}
url = "https://moneypuck.com/moneypuck/playerData/careers/gameByGame/regular/skaters/8479318.csv"
df <- read.csv(url)
```

## 1.3 - Wrangling data

### Removing variables

```{r}
df_all <- df[df$situation=="all",]
df_selected <- df_all[,c("gameDate", "opposingTeam", "I_F_goals", "I_F_points", "icetime", "gameScore")]
```

### Removing observations

```{r}
any(is.na(df_selected))
```

### Viewing data

```{r}
head(df_selected)
```

### Converting variables types

```{r}

```

### Creating derived variables

```{r}

```

## 1.4 - Variable breakdown

```{r}

```

## 1.5 - Exploratory data analysys

### Converting basic statistics 

```{r}
summary(df_selected)
sd(df_selected$I_F_goals)
sd(df_selected$I_F_points)
sd(df_selected$icetime)
sd(df_selected$gameScore)
```

### Returning data

```{r}
sqldf("SELECT min(gameScore), gameDate, opposingTeam FROM df_selected")
sqldf("SELECT max(gameScore), gameDate, opposingTeam FROM df_selected")
sqldf("SELECT max(I_F_goals), gameDate, opposingTeam FROM df_selected")
sqldf("SELECT max(I_F_points), gameDate, opposingTeam FROM df_selected")
```

### Computing and visualizing frequency distributions

```{r}
p_scores <- ggplot(df_selected, aes(x=gameScore)) +
  geom_histogram(fill="royalblue3", color="royalblue3", bins=10) +
  labs(title = "Game Score", 
       subtitle = "Career", 
       x = "Game Score",
       y = "Frequency") +
  theme(plot.title=element_text(face="bold"))

p_goals <- ggplot(df_selected, aes(x=I_F_goals)) +
  geom_histogram(fill="red3", color="red3", bins=5) +
  labs(title = "Goals", 
       subtitle = "Career", 
       x = "Goals",
       y = "Frequency") +
  theme(plot.title=element_text(face="bold")) 

p_points <- ggplot(df_selected, aes(x=I_F_points)) +
  geom_histogram(fill="springgreen3", color="springgreen3", bins=8) +
  labs(title = "Points", 
       subtitle = "Career", 
       x = "Game Score",
       y = "Frequency") +
  theme(plot.title=element_text(face="bold")) 

print(p_scores)
print(p_goals)
print(p_points)
                   
```

```{r}
sqldf("SELECT COUNT(*) FROM df_selected WHERE gameScore>=0")
sqldf("SELECT COUNT(*) FROM df_selected WHERE gameScore<0")
```

```{r}

```

### Computing and visualizing correlations

```{r}

```

### Computing and visualizing means and medians

```{r}

```

## 1.6 - Writing data

```{r}

```
